---
title: "Homework 3"
output:
  pdf_document: default
  html_document: default
date: 'Deadline: 9 February, 12:00'
subtitle: 'Linguistic Data: Quantitative Analysis and Visualisation'
---

The solutions should be submitted via GitHub.  

## Part 1. A preliminary training  

*Do not use R (RStudio) to solve problems in Part 1. Answers won't be evaluated. *
### Problem 1
Look at the following histogram and answer the questions.
```{r, echo=FALSE}
set.seed(1234)
x <- rnorm(20, mean=3, sd=6)
hist(x, col="tomato", main="")
```
a. What is the proportion of values in the sample that exceed $5$? Explain your answer.

5/20 or 1/4 of the sample exceeds value "5", because 2 columns to the right of "5" have frequency meanings 4+1=5 and other columns have frequency meanings 4+10+1=15, the total number of all frequency values is 20.

b. Indicate the interval where the median of this sample can lie. Explain your answer.

I guess, the median lies between x=-5 and x=0. Median is a line separating lower and higher parts of values, so it has to divide the biggest column and left and right parts of histogram.

c. How the histogram will change if we add an element $7$ to the sample? Explain your answer.

Тогда увеличится размер одного столбца с 5 до 10, так как у нас увеличится частота встречи значения, находящегося в том промежутке.
### Problem 2
a. Look at the histograms of two samples. They illustrate the distribution of normalized average reaction time to frequent words (in ms) in two groups of people.
```{r, echo=FALSE, fig.height=3}
set.seed(1234)
sample1 <- rnorm(100, 500, 50)
sample2 <- rnorm(100, 500, 20)
hist(sample1, col="lightblue", xlim=c(100, 900))
```
```{r, echo=FALSE, fig.height=3}
hist(sample2, col="lightgreen", xlim=c(100, 900))
```
Which of the samples has a larger variance? Explain your answer.

Это первая картинка, поскольку у нее более широкий диапазон значений x
\newpage
b. Look at the histograms of two samples.
```{r, echo=FALSE, fig.height=8}
set.seed(12)
sample1 <- rnorm(100, 3, 5)
sample2 <- rnorm(100, 10, 0.8)
par(mfrow=c(2,2))
hist(sample1, col="lightblue", xlim=c(-40, 40))
hist(sample2, col="lightgreen", xlim=c(5, 15))
```
Which of the samples has a larger variance? Explain your answer.

Это тоже первая картинка, поскольку здесь диапазон значений, на которые распространяется распределение, больше.

## Part 2
*Do not use R (RStudio) to solve problems in Part 2. Answers for problem 3 will be evaluated. Please paste YES or NO into (empty) code blocks and explain you answer below the block.*

### Problem 3

Below is the histogram of the number of mistakes students made while writing an examination essay in English. Look at the histogram and answer the questions.

```{r, echo=FALSE, fig.height=4}
set.seed(2)
mistakes <- rbeta(1000, 2, 7)*100
hist(mistakes, breaks=60, col= "deepskyblue", ylim=c(0, 40))
```

### 3.1
Is it true that $50$% students made more than $35$ mistakes?
```
No
```
Explain your answer below:

Частотности ошибок в количестве >= 35 в сумме своей меньше сумм частотностей ошибок, которые расположены слева от значения 35.

### 3.2
Is it true that most students made no more than $10$ mistakes?
```
No
```
Explain your answer below:

Не совсем, поскольку в значении 10, как мне кажется, не находится верхушка колпачка распределения.


### 3.3
Which of the following values is closer to be the median of `mistakes`: $10$, $20$, $30$, $40$?
```
20
```
Explain your answer below:

Как мне кажется, это 20, потому что медиана должна быть ровно посередине значений, и по моему мнению 20 находится ближе всего к верхушке колпачка распределений.

### Problem 4. Exact binomial test

The null hypothesis is that $p=0$ (i.e. no success is possible). In a dataset, there is only one success out of 1 000 000 observations. Will you reject the null hypothesis? 
```
Yes
```
Explain your answer below:

Вероятность успеха явно будет меньше альфы, равной 5%, и вообще она будет очень, очень маленькой (одна миллионная), поэтому я  отвергаю нулевую гипотезу.

## Part 3

*Use R (RStudio) to solve problems in Part 3. Your answers will be evaluated. Please paste R code into R code blocks and explain you answer below the block, if needed. *
### Problem 5 
Here is a sample of respondents' age:  
$44$, $50$, $42$, $64$, $66$, $42$, $72$, $56$, $72$, $54$, $46$, $48$, $48$, $52$, $50$, $66$, $84$.
### 5.1
Arrange them in a vector and call it `age`. 
```{r}
age <- c(44, 50, 42, 64, 66, 42, 72, 56, 72, 54, 46, 48, 48, 52, 50, 66, 84)
```
### 5.2
Examine the type of `age` variable (numeric, character, etc).
```{r}
class(age)
```
### 5.3
Plot the histogram of the vector `age` with $5$ bins. Change its color to any you want. (Use either R basic or ggplot2 style for plotting.)
```{r}
sample_size <- 5
hist(age, sample_size, col='lavender')
```
### Problem 6
Here is a series of words:  
*pie, bar, bar, pie, pie, bar, bar, chart*.

### 6.1
Arrange elements above in a vector and call it `words` 
```{r}
words <- c("pie", "bar", "bar", "pie", "pie", "bar", "bar", "chart")
```

### 6.2
Calculate the relative frequences of values in `words` measured in percent.
```{r}

#rel_freq = freq/words_size
(table(words)/sum(table(words)))*100 #домножим на 100 для процентов
```

## Problem 7. Position of verbs in verses

The dataset [“The last words in verses”](https://raw.githubusercontent.com/LingData2019/LingData2020/master/data/poetry_last_in_lines.csv) is based on texts written in the 1820s and 1920s (Corpus of Russian Poetry of the Russian National Corpus). Authors collected only one line per author to keep observations as independent from each other as possible. 

Variables:  
- Decade — decade of creation: 1820s, 1920s.  
- RhymedNwords — the number of words in the rhyming position (usually one word, but there are two words in cases such as _вина бы_ 'I would like to get) wine' (which is rhymed with _жабы_ 'toad', see http://russian-poetry.ru/Poem.php?PoemId=18261)).  
- RhymedNsyl — the number of syllables in the rhyming position.  
- UPoS — part of speech of the last word. 
- LineText — a sampled verse.  
- Author — the author of the text.  

Can we decide that in verses written in 1920s, verbs in the rhyming posision are used differently (more often or less often) than expected for verbs in general? 

Let's assume that the probability for verbs to be used in any position ('in general') is 17% (according to [](http://www.ruscorpora.ru/new/corpora-stat.html) ).

### 7.1 State hypothesis

What is your null hypothesis $H_0$ and what is the alternative hypothesis $H_1$?
```
H0: нет взаимосвязи, P=17%
H1: взаимосвязь есть, P>17%
```

### 7.2

Read the dataset [“The last words in verses”](https://raw.githubusercontent.com/LingData2019/LingData2020/master/data/poetry_last_in_lines.csv). Filter out the relevant observations from 1920s, calculate the number of verbs observed in the sample, and the sample size. 
```{r}
#install.packages("tidyverse")
library(tidyverse)
data <- read_tsv('https://raw.githubusercontent.com/LingData2019/LingData2020/master/data/poetry_last_in_lines.csv')
```
```{r}
#View(data)
```
### 7.3

Use an exact binomial test to calculate p-value. 
```{r}
data %>% filter(data$Decade=='1920s') -> data_filtered
data_filtered %>% filter(data_filtered$UPoS=='VERB') -> total_verbs

binom.test(nrow(total_verbs), nrow(data), p=0.17)
```

### 7.4 Interpret results

Give your interpretation of obtained p-value. Answer the initial question: Can we decide that in verses written in 1920s, verbs are used in the rhyming posision more often or less often than expected?
```
p-value получился очень маленьким, меньше 0,17, поэтому можно сказать, что взаимосвязь присутствует, но при этом мы говорим, что глаголы в целом использовались меньше.
```

### 7.5 
*(A bonus problem, extra points in evaluation)*.
Repeat 2.3 for verses written in the 1820s.
```{r}
data %>% filter(data$Decade=='1820s') -> data_filtered_18
data_filtered_18 %>% filter(data_filtered_18$UPoS=='VERB') -> total_verbs_18

binom.test(nrow(total_verbs_18), nrow(data), p=0.17)
```

Write down your general conlusions about data provided for both 1920s and 1820s data.
```
Так как p-value получился еще меньше, можем сделать выводы, что в 1920е глаголы использовались чаще, чем в 1820е.
```

### Problem 8. One-sample t-test
Using Icelandic data on vowel duration from seminar [Link](https://raw.githubusercontent.com/LingData2019/LingData2020/master/data/icelandic.csv) test the null hypothesis that the population mean of vowel duration in speaker `shg05` equals 73 (ms).
To perform a one-sample t-test, you can use the following example of R code: 
```
t.test(sample, mu = 7725) # mu is a population mean
```

### 8.1
Write down a two-tailed alternative hypothesis.
```
H1: Длится больше или меньше, чем 73
```

### 8.2
Perform a one-sample t-test. 
```{r}
data <- read_csv('https://raw.githubusercontent.com/LingData2019/LingData2020/master/data/icelandic.csv')
#View(data)

data %>% filter(data$speaker=='shg05') -> data_filtered
mean_vdur_shg05 <- mean(data_filtered$vowel.dur)
t.test(data_filtered$vowel.dur, mu = 73)
```

### 8.3
Interpret results. 
```
Так как p-value немного больше 5%, нулевую гипотезу всё равно не отвергаем
```

#### Supplementary reading
Use of exact binomial test in lingiistic research:  

* Gries, Stefan Th. "Phonological similarity in multi-word units." Cognitive Linguistics 22.3 (2011): 491-510. [Link](http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.186.7412&rep=rep1&type=pdf)  
Stefan Gries proves that alliteration is observed in multi-word expressions more often than in general.

* Harald Bayen (2008: 51-52) evaluates the probability of observing exactly one occurrence of the word _hare_ in the corpus sample of 1 mln words given its estimated frequency of 8.23 words per million according to the SELEX frequency database.

On measures of central tendency: 

* Levshina 2015, Chapter 3 (p. 48); Gries 2009, Chapter 1.3 (p. 116). 

On t-test: 

* Gries 2009, Chapter 3 (p. 198). 