---
title: 'HW 5: Correlations and linear models. Tests for categorial variables'
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(vcd) # for mosaic plots
```

## 1. Vowel reduction in Russian
Pavel Duryagin ran an experiment on perception of vowel reduction in Russian language. The dataset `shva` includes the following variables:  
_time1_ - reaction time 1  
_duration_ - duration of the vowel in the stimuly (in milliseconds, ms)  
_time2_ - reaction time 2  
_f1_, _f2_, _f3_ - the 1st, 2nd and 3rd formant of the vowel measured in Hz (for a short introduction into formants, see [here](https://home.cc.umanitoba.ca/~krussll/phonetics/acoustic/formants.html))  
_vowel_ - vowel classified according the 3-fold classification (_A_ - _a_ under stress, _a_ - _a/o_ as in the first syllable before the stressed one, _y_ (stands for shva) - _a/o_ as in the second etc. syllable before the stressed one or after the stressed syllable, cf. _g_[_y_]_g_[_a_]_t_[_A_]_l_[_y_] _gogotala_ `guffawed').  
In this part, we will ask you to analyse correlation between f1, f2, and duration.
The dataset is available [https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/data/duryagin_ReductionRussian.txt](here).
### 1.0 
Read the data from file to the variable `shva`.
```{r 1.0}
shva <- read_tsv(
  "https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/data/duryagin_ReductionRussian.txt")
shva
```

### 1.1 
Scatterplot `f1` and `f2` using `ggplot()`. 
Design it to look like the [following](https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/img/duryagin1.png).
```{r 1.1}
shva %>%
  ggplot(aes(f2, f1, colour = vowel)) +
  geom_point() +
  scale_x_reverse() +
  scale_y_reverse()
```

### 1.2 
Plot the boxplots of `f1` and `f2` for each vowel using `ggplot()`.
Design it to look like [this](https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/img/duryagin2.png) and [this](https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/img/duryagin3.png).
```{r 1.2}
# f1 boxplot
shva %>%
  ggplot(aes(vowel, f1, fill=vowel)) +
  coord_flip() +
  geom_boxplot()

# f2 boxplot
shva %>%
  ggplot(aes(vowel, f2, fill=vowel)) +
  coord_flip() +
  geom_boxplot()
```

### 1.3 
Calculate Pearson's correlation of `f1` and `f2` (all data)
```{r 1.3}
cor.test(shva$f1, shva$f2, method = 'pearson')
```

### 
1.4 Calculate Pearson's correlation of `f1` and `f2` for each vowel
```{r 1.4}
shva %>% filter(vowel == "A") -> shva_A
cor.test(shva_A$f1, shva_A$f2, method = 'pearson')
```

```{r 1.4}
shva %>% filter(vowel == "a") -> shva_a
cor.test(shva_a$f1, shva_a$f2, method = 'pearson')
```

```{r 1.4}
shva %>% filter(vowel == "y") -> shva_y
cor.test(shva_y$f1, shva_y$f2, method = 'pearson')
```

## 2 Linear regressions

### 2.1.1 
Use the linear regression model to predict `f2` by `f1`.
```{r 2.1.1}
lr <- lm(data = shva, f2 ~ f1)
summary(lr)
```

### 2.1.2 
Write down the equation for f2 using coefficients from the model (e.g. $y =  b + kx$)
```{r 2.1.2}
f2 = 1639.70215 - 0.42875*f1
```

### 2.1.3 
Provide the adjusted R$^2$
```{r 2.1.3}
summary(lr)$adj.r.squared
```

### 2.1.4 
Add the regression line in the scatterplot 1.1.
```{r 2.1.4}
shva %>%
  ggplot(aes(f2, f1, colour = vowel)) +
  geom_point() +
  scale_x_reverse() +
  scale_y_reverse() + 
  geom_smooth(method = lm)
```

### 2.1.5 
Make a scatter plot for `f1` and `f2` grouped by vowels. 
Use `ggplot()` and `facet_wrap()`.

```{r 2.1.5}
ggplot(shva, aes(y = f1, x = f2)) +
  geom_point(aes(color = vowel)) +
  scale_x_reverse() +
  scale_y_reverse() +
  facet_wrap(~ vowel)
```

### 2.2.1 
Use the linear regression model to predict `f2` by `f1` and `vowel`.
```{r 2.2.1}
multi_lr <- lm(data = shva, f2 ~ f1 + vowel)
summary(multi_lr)
```

### 2.2.2 
What is the intercept of the model?
```{r 2.2.2}
multi_lr$coefficients[1]
```

### 2.2.3 
Provide the adjusted R$^2$
```{r 2.2.3}
summary(multi_lr)$adj.r.squared
```

### 2.2.4 
Write down your general conclusions about the relationship between `f1`, `f2`, and `vowels`.

```
f1 и f2 связаны следующим образом: если мы рассмотрим f2 + f1 и f2 + f1 и vowels, то во втором варианте коэффициенты будут больше, что значит следующее: f2 зависит от vowels.
```

## 3. Dutch causative constructions

When the Dutch use two near-synonymous periphrastic causative verbs, *doen* and *laten*?
```
       De politie deed/liet de auto stoppen.
  lit. the police did/let the car stop 
       'The police stopped the car'
```

This is a data set on two rival constructions with *doen* and *laten* sampled from the newspaper corpora. The data frame includes 500 observations on the following 7 variables:

* `Aux` -- verb: a factor with levels `doen` and `laten`  
* `CrSem` -- the semantic class of the Causer: a factor with levels `Anim` (animate) and `Inanim` (inanimate)  
* `CeSem` -- the semantic class of the Causee: a factor with levels `Anim` (animate) and `Inanim` (inanimate)  
* `CdEvSem` -- the semantic domain of the caused event expressed by the Effected Predicate: a factor with levels `Ment` (mental) and `NonMent` (e.g. physical or social)  
* `CeSynt` -- the syntactic status of the Causee: a factor with levels `Clause`, `Impl` (implicit, not expressed), `NP` (noun phrase), `PP` (prepositional phrase)    
* `EPTrans` -- transitivity or intransitivity of the effected predicate, a factor with two levels `Tr` and `Intr`  
* `Country` -- a factor with levels `BE` (Belgium) and `NL` (Netherlands)  
* `Domain` -- a factor with four levels for newspaper domains.    

Data from Natalya Levshina's `RLing` package available (here)[https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/data/dutch_causatives.csv]
Read more on the constructions in [Levhina, Geerarts, Speelman 2014](https://www.academia.edu/7287585/Dutch_causative_constructions_Quantification_of_meaning_and_meaning_of_quantification_with_Dirk_Geeraerts_and_Dirk_Speelman_).

### 3.0 
Read the data from file to the variable `d_caus`.
```{r 3.0}
d_caus <- read.csv("https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/data/dutch_causatives.csv")
summary(d_caus)
```

### 3.1 
We are going to test whether the association between `Aux` and other categorical variables (`Aux` ~ `CrSem`, `Aux` ~ `CeSem`, etc) is statistically significant. The assiciation with which variable should be analysed using Fisher's Exact Test and not using Pearson's Chi-squared Test? Is this association statistically significant?

Можно применить хи-квадрат только с CeSynt, поскольку только у нее ожидаемая частота больше 5.

```{r 3.1}
table(d_caus$Aux, d_caus$CrSem)
table(d_caus$Aux, d_caus$CeSem)
table(d_caus$Aux, d_caus$CdEvSem)
table(d_caus$Aux, d_caus$CeSynt)
table(d_caus$Aux, d_caus$EPTrans)
table(d_caus$Aux, d_caus$Country)
table(d_caus$Aux, d_caus$Domain)
fisher.test(d_caus$Aux, d_caus$CeSynt)
```

### 3.2. 
Test the hypothesis that `Aux` and `EPTrans` are not independent with the help of Pearson's Chi-squared Test. 
```{r 3.2}
aux_ept <- chisq.test(d_caus$Aux, d_caus$EPTrans)
aux_ept
```
 p-value < 0.05, следовательно, связь между Aux и  CeSynt статистически значима

### 3.3 
Provide expected frequencies for Pearson's Chi-squared Test of `Aux` and `EPTrans` variables.
```{r 3.3}
aux_ept$expected
```
p-value < 0.05, следовательно, есть связь между переменными

### 3.4. 
Calculate the odds ratio for observed frequencies of `Aux` and `EPTrans`
For 2×2 contigency table
$$\begin{matrix}
a & b \\
c & d
\end{matrix}
$$
one can find *odds ratio* as $(a/c)/(b/d)$.

```{r 3.4}
abcd <- aux_ept$observed
abcd
odds_ratio = (abcd[1, 1]/abcd[2, 1])/(abcd[1, 2]/abcd[2, 2])
odds_ratio
```

### 3.4.1 
Find odds ratio for expected frequencies of `Aux` and `EPTrans`

```{r 3.4.1}
abcd <- aux_ept$expected
abcd
odds_ratio = (abcd[1, 1]/abcd[2, 1])/(abcd[1, 2]/abcd[2, 2])
odds_ratio
```
What can you say about odds ratio of expected frequencies for arbitrary data?

Так как odds ratio = 1, данные считаются независимыми.


### 3.5 
Calculate effect size for this test using Cramer's V (phi).
```{r 3.5}
library(lsr)
cramersV(aux_ept$observed)
```

### 3.6. 
Report the results of independence test using the following template:
```
We have / have not found a significant association between variables ... and ... (p < 0.001).  The odds of ... were ... times higher / lower in (group ....) than in (group ....). Effect size is large / medium / small (Cramer's V = ....).
```
We have found a significant association between variables Aux and EPTrans (p < 0.001). The odds of "doen" were 2.6 times higher in intransitive than in transitive. Effect size is small (Cramer's V = 0.17).

### 3.7 
Visualize the distribution using mosaic plot.
Use `mosaic()` function from `vcd` library.
```{r 3.7}
mosaic(~ Aux + EPTrans, data=d_caus, shade=TRUE, legend=TRUE)
```

Below is an example of how to use mosaic() with three variables.
```{r 3.7.1}
mosaic(~ Aux + CrSem + Country, data=d_caus, shade=TRUE, legend=TRUE)
```

### 3.8 
Why is it not recommended to run multiple Chisq tests of independence on different variables within your dataset whithout adjusting for the multiplicity? (i.e. just testing all the pairs of variables one by one)  
```
Если мы будем тестировать гипотезы отдельно, то вероятность получить ошибку будет расти.
```

### 3.9 
Provide a short text (300 words) describing the hypothesis of this study and the results of your analysis.

```
Мы изучали связь между вспомогательными глаголами doen и laten и переходностью в голландском языке. Гипотеза исследования заключается в том, что выбор глагола зависит от его переходности. После того, как мы применили тест хи-квадрат, мы выяснили, что существует статистически значимая связь между переходностью и вспомогательными глаголами.
```